---
title: "Testdoc_pivot_2024"
author: "Dr. Pivot"
format: 
  docx:
    reference-doc: template1.docx
    toc: true
    toc-depth: 3
execute: 
  echo: true
  warning: false
  output: asis
fig.dpi: 300
fig-width: 6
fig-height: 4
editor: visual
---

## How to use pivot function

#### STEP 1: load packages and read the data file. The data file (widedata.xlsx) has the variable name "rawdata"

```{r}
pacman::p_load(conflicted, tidyverse, readxl, wrappedtools, here, flextable)
rawdata <- read_excel(here("~/Desktop/2024_Bioinfo_Biostats/CQ/Bioinfo_Biostat_Weiterbildung/Kurs/Module4_Applied_Biostatistics/RStatsbook2/widedata.xlsx"))
```

#### Visualizing the content of rawdata

```{r}
rawdata
```

\newpage

#### Visualizing the rawdata using flextable()

```{r}
rawdata |>
  flextable()|>
  theme_zebra() |>
  set_table_properties(width =1,
                       layout = "autofit")
```

#### STEP 2: Modify table rawdata to longer using pivot_longer

##### Once the rawdata is in Global environment, start using pivot_longer to modify the table. This modification doesn't modify the original rawdata. Remember: to export the table with a comprehensive aesthetics, please use flextable() function.

```{r}
rd_long1 <- pivot_longer(rawdata,
                         starts_with("We"))
rd_long1 |>
  flextable()|>
  theme_zebra() |>
  set_table_properties(width =1,
                       layout = "autofit")
  
```

\newpage

#### STEP 3: Improve a bit more the pivot_longer

```{r}
rd_long2 <- pivot_longer(rawdata,
                         starts_with("We"),
                         values_to="Weight [g]")
rd_long2|>
  flextable()|>
  theme_zebra() |>
  set_table_properties(width =1,
                       layout = "autofit")
```

\newpage

#### STEP 4: Improve a bit more the pivot_longer

```{r}
rd_long3 <- pivot_longer(rawdata,
                         starts_with("We"), # here 'cols' was omitted but by following the order of factors, we can omit it
                         values_to="Weight [g]",
                         names_to="Time [h]",
                         names_pattern = ".+ (\\d+).+") #regular expression meaning: search for one or more than one character, followed by a space and then for one or more numbers, and after the number there's something else
rd_long3|>
  flextable()|>
  theme_zebra() |>
  set_table_properties(width =1,
                       layout = "autofit")
```

#### STEP 5: from longer to wider

```{r}
rd_long <-
  pivot_longer(rawdata,
               cols = starts_with("We"),
               names_to=c(".value","Time [h]"),
               names_pattern = "(W.+\\]) (\\d+).+") # dot(.) is a wild card and it can be anything. Plus (+) means that there's one or more characters after "W"

rd_wide <-
  pivot_wider(rd_long,
              names_from = `Time [h]`,
              values_from = `Weight [g]`,
              names_glue = "Weight [g] @{`Time [h]`}h") # in {} we are referring the name of the columns generated from "Time [h]"
rd_wide|>
  flextable()|>
  theme_zebra() |>
  set_table_properties(width =1,
                       layout = "autofit")

```

#### STEP 6: 

#### Changing the appearance 1

```{r}
rd_wide|>
  flextable()|>
  theme_zebra(odd_header = "#ccd5ae") |>
  set_table_properties(width =1,
                       layout = "autofit")
```

#### STEP 6:

#### Changing the appearance 2

```{r}
rd_wide|>
  flextable()|>
  theme_tron_legacy() |>
  set_table_properties(width =1,
                       layout = "autofit")
```
